---
layout: post
title: "Hyperspectral Remote Sensing"
date:   2015-10-22 20:49:52
authors: "Keely Roth, Mike Alonzo, Jonah Duckles"
dateCreated:  2015-10-22 20:49:52
lastModified: 2015-10-22 14:28:52
category: remote-sensing-workshop
tags: [module-3]
mainTag: GIS-Remote-Sensing-Data
description: "Overview of hyperspectral analysis in R"
code1:
image:
  feature: lidar_GrandMesa.png
  credit: LiDAR data collected over Grand Mesa, Colorado - National Ecological Observatory Network (NEON)
  creditlink: http://www.neoninc.org
permalink: /R/Raster-Data-In-R/
code1: /R/2015-07-22-Introduction-to-Raster-Data-In-R.R
comments: true

---

<section id="table-of-contents" class="toc">
  <header>
    <h3>Contents</h3>
  </header>
<div id="drawer" markdown="1">
*  Auto generated table of contents
{:toc}
</div>
</section><!-- /#table-of-contents -->



## About

## R Skills Level


## Goals and Objectives
* understanding of spectral concepts: bands and wavelengths
* open & display multispectral and hyperspectral data
* geographically clip raster imagery
* plot and compare single-pixel spectra
* Understanding of information content of visible, multispectral and hyperspectral data
* recognize general spectral signatures
* calculate and understand the physical basis for spectral metrics
* create and threshold based on histogram visualizations of spectral information
* standard deviation moving window

## Things You'll Need to Complete this Lesson

### R libraries
* `raster`
* `rgdal`
* `rhdf5`
* `sp`
* `manipulate`
* `plyr`*


``` {r}
install.packages('raster')
install.packages('rgdal')
install.packages('rhdf5')
install.packages('sp')
install.packages('manipulate')
install.packages('plyr')

library('raster')
library('rgdal')
library('rhdf5')
library('sp')
library('manipulate')
library('plyr')
```

### Tools

* R
* RStudio

### Data
* Visible High Res 10cm resolution
* Hyperspectral HDF5 1m resolution
* Multispectral Data ~30m resolution

``` {r}
#download and save each datasets
```


## Recommended Pre Lesson Reading

## Notes for Formatting Lessons
fill in outline for each module (think about story arc):
with lesson obj
lesson text/concept blocks
chunk in programming steps
Challenges
organization:
Objective
Tutorial
Functions
Challenge





## NEON Data Resolutions
|             |  Spatial             | Spectral         | Temporal |
|-------------|----------------------|------------------|----------|
|NEON Visible       | Very High (eg 10cm)  | 3 broad bands    | 1-2/year|
|Landsat Multispectral | Moderate (eg 30m)    | 6 broad bands    | ~22/year|
|NEON hyperspectral | High (eg 1m)         | 424 FIXME narrow bands| 1-2/year|

## Lesson 1: Opening and Exploring Multispectral and Hyperspectral Imagery

Objective is XXX.

To get started, we'll explore each of the three image datasets used in this module.
First we'll open and 

``` {r}
# Open RGB camera image
# Display single bands 
# Create RGB stack
# Display RGB stack
```
Challenge:

## Lesson 2: Spectral Signatures

Here we'll learn about the spectrum and compare the information content in multi-v-hyper

``` {r}
#code goes here
```

## Lesson 3: Spectral Metrics

Now that we understand the driving forces behind the spectrum, we can explore different ways of capturing that info.

``` {r}
#code goes here
```

## Lesson 4: Mapping Spectral Metrics

Now we'll make a map of these metrics and use it to answer some questions.

``` {r}
#code goes here
```


###  viz
Open
Display one band of image (for each band)
Create stack and display rgb composite
Open vector and overlay
Clip raster using the vector (using crop?)

```{r}
#load 10 cm vis dataset

```

Use image(red) image(green) image(blue)
SJER_camera.tif
import(raster)
plotrgb()


### multi
Open the geotiffs
Display one band of image
Create 6-band stack and display a true color composite
Open vector and overlay
Clip stack using the vector (using crop?)

```{r}
#load landsat subset here
```

###hyper
Open HDF5
(maybe look at some metadata)
Slice out one band of image
Find roughly the same wavelenght as done above 
Slice out true color wavelengths
Create 3-band stack and display
Open vector and overlay
Clip the full set of bands using the vector (using crop?)



```{r}
#load hyperspectral data as h5



```



### Compare
3-by raster plot
3-by spectral plot

## Clip

Why do we clip?

Apply a clip across all three rasters to about two landsat pixels

## Understanding hyperspectral signatures
Physics of light and reflectance

Static diagram with spectral regions

Fire trucks (that are yellow/green)!

### Plot example spectral signatures

* Plot some spectral signatures

### Compare spectral signatures
* Compare
* plot of several materials (trees, dirt, grass, water)

### Spectral metrics
* lines at 550 for green leaves
* lines at 684 for chlorophyl absorption
* Calculate spectral metrics
  * Naroww band NDVI
  * Water metric
  * Stress metric
* Single pixel calculation
* Creating a new raster based on the metrics with histogram
* Threshold the new computed raster



##About
This activity will walk you through the fundamental principles of working
with raster data in R.

**R Skill Level:** Intermediate - you've got the basics of `R` down.

<div id="objectives">

<h3>Goals / Objectives</h3>
After completing this activity, you will know:
<ol>
<li>What a raster dataset is and its fundamental attributes.</li>
<li>How to import rasters into `R` using the raster library.</li>
<li>How to perform raster calculations in `R`.</li>
</ol>

<h3>Things You'll Need To Complete This Lesson</h3>

<h3>R Libraries to Install:</h3>
<ul>
<li><strong>raster:</strong> <code> install.packages("raster")</code></li>
<li><strong>rgdal:</strong> <code> install.packages("rgdal")</code></li>

</ul>
<h4>Tools To Install</h4>

Please be sure you have the most current version of `R` and preferably
R studio to write your code.


<h4>Data to Download</h4>

Download the raster and *in situ* collected vegetation structure data:
<ul>
<li><a href="http://neonhighered.org/Data/LidarActivity/CHM_InSitu_Data.zip" class="btn btn-success"> DOWNLOAD Sample NEON LiDAR data in Raster Format & Vegetation Sampling Data</a></li>
<li><a href="{{ site.baseurl }}/data/rasterLayers_tif.zip" class="btn btn-success"> DOWNLOAD NEON imagery data (tiff format) California Domain 17 (D17)</a></li>
</ul>

<p>The LiDAR and imagery data used to create the rasters in this dataset were
collected over the San Joaquin field site located in California (NEON Domain 17)
and processed at <a href="http://www.neoninc.org" target="_blank" >NEON </a>
headquarters. The entire dataset can be accessed by request from the NEON website.</p>  

<h4>Recommended Pre-Lesson Reading</h4>
<ul>
<li>
<a href="{{ site.baseurl }}/GIS-Spatial-Data/Working-With-Rasters/" target="_blank">
The Relationship Between Raster Resolution, Spatial extent & Number of Pixels - in R</a>
</li>
<li>
<a href="http://cran.r-project.org/web/packages/raster/raster.pdf" target="_blank">
Read more about the `raster` package in R.</a>
</li>
</ul>
</div>

#About Raster Data
Raster or "gridded" data are data that are saved in pixels. In the spatial world,
each pixel represents an area on the Earth's surface. For example in the raster
below, each pixel represents a particular land cover class that would be found in
that location in the real world.
<a href="{{ site.baseurl }}/GIS-Spatial-Data/Working-With-Rasters/"> More on
rasters here</a>.

```{r chunk-1}
#code will go here!

```
